The title

### 2024-10-04 05_Manuscript/manuscript.Rmd
updating data screening and efa

```{r libraries}
library(rio)
library(dplyr)
set.seed(8943)
library(mice)

@@

```{r load-data, eval = T}
sona <- import("../03_Data/sona_data_deidentify.csv") %>% 
  mutate(where = "sona")
mturk <- import("../03_Data/mturk_data_deidentify.csv") %>% 
  mutate(where = "mturk")
```

@@

```{r screen-sona-accuracy-out, eval = F}
# Accuracy ----
summary(sona %>% 
          select(Q2_1:Q2_21, Q4_1:Q4_20, Q5_1:Q5_20))
# everything is within the expected range

apply(sona %>% 
          select(Q2_1:Q2_21, Q4_1:Q4_20, Q5_1:Q5_20), 
      2, mean, na.rm = TRUE)
apply(sona %>% 
          select(Q2_1:Q2_21, Q4_1:Q4_20, Q5_1:Q5_20), 
      2, sd, na.rm = TRUE)
#

@@

noout = subset(nolowqual, mahal < cutoff)
#42 participants excluded as outliers
```

-   try running with outliers as well --\> if differences: then talk
    about what is going on with the outlier people

# Assumptions -------------------------------------------------------------

# Linearity

random = rchisq(nrow(noout[ , 37:97]), 7) fake = lm(random\~.,
data=noout[ , 37:97])

standardized = rstudent(fake) {qqnorm(standardized) abline(0,1)} #seems
okay

# Normality

hist(standardized, breaks=15) #a little skewed but mostly centered on 0
and between -2 and 2 so okay

# Homog/s

fitvalues = scale(fake\$fitted.values) {plot(fitvalues, standardized)
abline(0,0) abline(v = 0)} #some homogeneity issues but we'll accept it

write.csv(noout, file = "sona_data_screened.csv", row.names = F)
library(beepr) beep(sound = 5)

\`\`\`

>>>>

```{r libraries}
library(rio)
library(dplyr)
set.seed(8943)
library(mice)
library(psych)

@@

```{r load-data, eval = T}
sona <- import("../03_Data/sona_data_deidentify.csv") %>% 
  mutate(where = "sona")
sona_replaced <- import("../03_Data/sona_data_replaced.csv")
sona_page1 <- import("../03_Data/sona_page1_sad.csv")
sona_page2 <- import("../03_Data/sona_page2_sad.csv")
sona_page3 <- import("../03_Data/sona_page3_sad.csv")
sona_good <- import("../03_Data/sona_data_goodqual.csv")
sona_noout <- import("../03_Data/sona_data_noout.csv")

mturk <- import("../03_Data/mturk_data_deidentify.csv") %>% 
  mutate(where = "mturk")
mturk_replaced <- import("../03_Data/mturk_data_replaced.csv")
mturk_page1 <- import("../03_Data/mturk_page1_sad.csv")
mturk_page2 <- import("../03_Data/mturk_page2_sad.csv")
mturk_page3 <- import("../03_Data/mturk_page3_sad.csv")
mturk_good <- import("../03_Data/mturk_data_goodqual.csv")
mturk_noout <- import("../03_Data/mturk_data_noout.csv")

```

@@

```{r screen-sona-accuracy-out, eval = F}
# Accuracy ----
summary(sona %>% 
          select(Q2_1:Q2_21, Q4_1:Q4_20, Q5_1:Q5_20))
apply(sona %>% 
          select(Q2_1:Q2_21, Q4_1:Q4_20, Q5_1:Q5_20), 
      2, mean, na.rm = TRUE)
apply(sona %>% 
          select(Q2_1:Q2_21, Q4_1:Q4_20, Q5_1:Q5_20), 
      2, sd, na.rm = TRUE)
# everything is within the expected range

@@

noout <- subset(nolowqual, mahal < cutoff)
write.csv(noout, "../03_Data/sona_data_noout.csv", row.names = F)
```

```{r sona-assumptions, eval = F} 
random <- rchisq(nrow(sona_good), 7) 
fake <- lm(random ~ ., data = sona_good %>% 
             select(Q2_1:Q5_20)  %>% 
             select(-Q2_21, -Q4_20, -Q5_16))
standardized <- rstudent(fake)
fitvalues <- scale(fake$fitted.values)

# Linearity ----
{
  qqnorm(standardized)
  abline(0,1)
}

# Normality -----
hist(standardized, breaks=15)

# Homoscedasticity -----
{
  plot(fitvalues, standardized)
  abline(0,0) 
  abline(v = 0)
}

random <- rchisq(nrow(sona_noout), 7) 
fake <- lm(random ~ ., data = sona_noout %>% 
             select(Q2_1:Q5_20)  %>% 
             select(-Q2_21, -Q4_20, -Q5_16))
standardized <- rstudent(fake)
fitvalues <- scale(fake$fitted.values)

# Linearity -----
{
  qqnorm(standardized)
  abline(0,1)
}

# Normality -----
hist(standardized, breaks=15)

# Homoscedasticity -----
{
  plot(fitvalues, standardized)
  abline(0,0) 
  abline(v = 0)
} 
```

```{r screen-mturk-accuracy-out, eval = F}
# Accuracy ----
summary(mturk %>% 
          select(Q2_1:Q2_21, Q4_1:Q4_20, Q5_1:Q5_20))
apply(mturk %>% 
          select(Q2_1:Q2_21, Q4_1:Q4_20, Q5_1:Q5_20), 
      2, mean, na.rm = TRUE)
apply(mturk %>% 
          select(Q2_1:Q2_21, Q4_1:Q4_20, Q5_1:Q5_20), 
      2, sd, na.rm = TRUE)
# everything is within the expected range

# rows
missing <- apply(mturk %>% 
          select(Q2_1:Q2_21, Q4_1:Q4_20, Q5_1:Q5_20), 
          1, 
          percentmiss) 
table(missing)
replacepeople <- subset(mturk, missing <= 5)
dontpeople <- subset(mturk, missing > 5)

# putting data back together
# excluding participants with > 5% missing (7 total)
allcolumns <- cbind(dontcolumn, nomiss)
summary(allcolumns)
write.csv(allcolumns, "../03_Data/mturk_data_replaced.csv", row.names = F)
```

```{r screen-mturk-SAD, eval = F}
# SAD 
nomissing <- import("../03_Data/mturk_data_replaced.csv")

# Block 1
page1 <- SAD(dat = nomissing %>% 
               select(Q2_1:Q2_21),
    rt = nomissing$`Q10_Page Submit`,
    min = 1, 
    max = 5, 
    partno = nomissing$ResponseId, 
    click = nomissing$`Q10_Click Count`, 
    manvec = nomissing$Q2_21, 
    mancor = 1, 
    char = 1626)
write.csv(page1, "../03_Data/mturk_page1_sad.csv", row.names = F)

# Block 2
page2 <- SAD(dat = nomissing %>% 
               select(Q4_1:Q4_20), 
            rt = nomissing$`Q11_Page Submit`,
            min = 1, 
            max = 5, 
            partno = nomissing$ResponseId, 
            click = nomissing$`Q11_Click Count`, 
            manvec = nomissing$Q4_20, 
            mancor = 4, 
            char = 1491)
write.csv(page2, "../03_Data/mturk_page2_sad.csv", row.names = F)

# Block 3
page3 <- SAD(dat = nomissing %>% 
               select(Q5_1:Q5_20), 
            rt = nomissing$`Q12_Page Submit`,
            min = 1, 
            max = 5, 
            partno = nomissing$ResponseId, 
            click = nomissing$`Q12_Click Count`, 
            manvec = nomissing$Q5_16, 
            mancor = 3, 
            char = 1496)
write.csv(page3, "../03_Data/mturk_page3_sad.csv", row.names = F)

nolowqual <- subset(nomissing, totalbadUP < 4)
write.csv(nolowqual, "../03_Data/mturk_data_goodqual.csv", row.names = F)
```

```{r mturk-outliers, eval = F}
nolowqual <- import("../03_Data/mturk_data_goodqual.csv")

noout <- subset(nolowqual, mahal < cutoff)
write.csv(noout, "../03_Data/mturk_data_noout.csv", row.names = F)
```

```{r mturk-assumptions, eval = F} 
random <- rchisq(nrow(mturk_good), 7) 
fake <- lm(random ~ ., data = mturk_good %>% 
             select(Q2_1:Q5_20)  %>% 
             select(-Q2_21, -Q4_20, -Q5_16))
standardized <- rstudent(fake)
fitvalues <- scale(fake$fitted.values)

# Linearity ----
{
  qqnorm(standardized)
  abline(0,1)
}

# Normality -----
hist(standardized, breaks=15)

# Homoscedasticity -----
{
  plot(fitvalues, standardized)
  abline(0,0) 
  abline(v = 0)
}

random <- rchisq(nrow(mturk_noout), 7) 
fake <- lm(random ~ ., data = mturk_noout %>% 
             select(Q2_1:Q5_20)  %>% 
             select(-Q2_21, -Q4_20, -Q5_16))
standardized <- rstudent(fake)
fitvalues <- scale(fake$fitted.values)

# Linearity -----
{
  qqnorm(standardized)
  abline(0,1)
}

# Normality -----
hist(standardized, breaks=15)

# Homoscedasticity -----
{
  plot(fitvalues, standardized)
  abline(0,0) 
  abline(v = 0)
} 
```

```{r data-combine-split}
combo <- bind_rows(sona_good, mturk_good)
combo_noout <- bind_rows(sona_noout, mturk_noout)

random_split <- c(
  sample(sona_good$ResponseId, nrow(sona_good)/2),
  sample(mturk_good$ResponseId, round(nrow(mturk_good)/2))
)

efaDF <- combo %>% 
  filter(ResponseId %in% random_split) %>% 
  select(Q2_1:Q5_20)  %>% 
  select(-Q2_21, -Q4_20, -Q5_16)
cfaDF <- combo %>% 
  filter(!(ResponseId %in% random_split)) %>% 
  select(Q2_1:Q5_20)  %>% 
  select(-Q2_21, -Q4_20, -Q5_16)

nrow(efaDF)
nrow(cfaDF)

efaDF_noout <- combo_noout %>% 
  filter(ResponseId %in% random_split) %>% 
  select(Q2_1:Q5_20)  %>% 
  select(-Q2_21, -Q4_20, -Q5_16)
cfaDF_noout <- combo_noout %>% 
  filter(!(ResponseId %in% random_split)) %>% 
  select(Q2_1:Q5_20)  %>% 
  select(-Q2_21, -Q4_20, -Q5_16)

nrow(efaDF_noout)
nrow(cfaDF_noout)
```

## EFA

### Number of Factors

```{r}
number_items <- fa.parallel(efaDF, #data frame
                            fm="ml", #math
                            fa="fa") #only efa
sum(number_items$fa.values > 1)
sum(number_items$fa.values > .7)
```

### Simple Structure

```{r}
EFA_fit <- fa(efaDF, #data
              nfactors = 2, #number of factors
              rotate = "oblimin", #rotation
              fm = "ml") #math

EFA_fit

efa_loadings <- as.data.frame(unclass(EFA_fit$loadings)) %>% 
  mutate(keep = ifelse(
    ML1 >= .30, "yes", ifelse(
      ML2 >= .30, "yes", ifelse(
        ML1 >= .30 & ML2 >= .30, "no", "no"
      )
    )
  )) %>% 
  mutate(question = rownames(.))

EFA_fit_2 <- fa(efaDF %>% 
                  select(efa_loadings %>% 
                           filter(keep == "yes") %>% 
                           pull(question)), #data
              nfactors = 2, #number of factors
              rotate = "oblimin", #rotation
              fm = "ml") #math

EFA_fit_2

efa_loadings_2 <- as.data.frame(unclass(EFA_fit_2$loadings)) %>% 
  mutate(keep = ifelse(
    ML1 >= .30, "yes", ifelse(
      ML2 >= .30, "yes", ifelse(
        ML1 >= .30 & ML2 >= .30, "no", "no"
      )
    )
  )) %>% 
  mutate(question = rownames(.))
```

```{r}
EFA_fit <- fa(efaDF, #data
              nfactors = 3, #number of factors
              rotate = "oblimin", #rotation
              fm = "ml") #math

EFA_fit

efa_loadings <- as.data.frame(unclass(EFA_fit$loadings)) %>% 
  mutate(keep = ifelse(
    ML1 >= .30, "yes", ifelse(
      ML2 >= .30, "yes", ifelse(
        ML3 >= .30, "yes", "no"
        )
      )
    )
  ) %>% 
  mutate(question = rownames(.)) %>% 
  mutate(
    keep = ifelse(
          (ML1 >= .30 & ML2 >= .30) | 
            (ML1 >= .30 & ML3 >= .30) |
            (ML3 >= .30 & ML2 >= .30), "no", keep
          )
  )

EFA_fit_2 <- fa(efaDF %>% 
                  select(efa_loadings %>% 
                           filter(keep == "yes") %>% 
                           pull(question)), #data
              nfactors = 3, #number of factors
              rotate = "oblimin", #rotation
              fm = "ml") #math

EFA_fit_2

efa_loadings_2 <- as.data.frame(unclass(EFA_fit_2$loadings)) %>% 
  mutate(keep = ifelse(
    ML1 >= .30, "yes", ifelse(
      ML2 >= .30, "yes", ifelse(
        ML3 >= .30, "yes", "no"
        )
      )
    )
  ) %>% 
  mutate(question = rownames(.)) %>% 
  mutate(
    keep = ifelse(
          (ML1 >= .30 & ML2 >= .30) | 
            (ML1 >= .30 & ML3 >= .30) |
            (ML3 >= .30 & ML2 >= .30), "no", keep
          )
  )
```
